Consistent Polyhedral Surrogates via Embedding

### 2021-08-04 notes/top-k.tex
Finishing proof

	(c) Cannot happen, since $H_p = \emptyset \implies B_p = \{i \mid p_i \in \left(\frac{1}{k+1}, \frac{1}{k+1}\right]\} = \emptyset$.

	(d) Since $p_H > \frac H k$, we have $p_M < 1 - H_k$.
	\jessie{TODO}

	(e) Suppose $p_H = 1 - \epsilon$.
	For all $b \in B_p$, we then have $p_b > \frac{k \epsilon}{(k+1)(k-H)}$.
	By simplex constraints, we then have 
	\begin{align*}
	1 - \epsilon + \frac{Bk\epsilon}{(k+1)(k-H)}&\leq 1 \\
	\iff \frac{Bk}{(k+1)(k-H)}&\leq 1 \\
	Bk &\leq (k+1)(k-H)\\
	B &\leq \frac{k^2 + k - H - Hk}{k}\\
	B &\leq k+1-H - \frac H k
	&< k+1 - H
	\end{align*}
	Thus, $B+H < k+1$.

@@

	(g)

	\jessie{Thought: Break into cases: the nonzero elements in (a) H (b) M (c) B (d) H,M (e) H,B (f) B,M (g) H,M,B.  Turns out that a,b are pretty trivial, c and f can never happen.  H,M has an untyped proof in my notes, but e and g have yet to be shown.  There's a tension between modifying the sets so this proof is more immediate and the result being true.  I think we can determine the sets algorithmically so this proof is more immediate, but showing it is a correct property value might be harder.}

@@

	Now for any sets $H,M$ such that $H, M \subseteq [n]$, $H \cap M = \emptyset$, $|H| + |M| \leq k$, and $|H| < k$ with $k <n$, consider the distribution $p_{HM} = (\frac{1}{k} + \varepsilon)\ones_{H} + (\frac{1}{k+1} + \varepsilon)\ones_M + (1 - ((\frac{|H|}{k} + |H| \varepsilon) + (\frac{|M|}{k+1} + |M| \varepsilon))) \ones_{[n] \setminus H\cup M} $.
\end{proof}

@@

\begin{lemma}\label{lem:hat-gamma-opt-one-dir}
	Fix $p \in \simplex$.
	For small enough $\varepsilon$ and any $\hat y \in \Y$, consider $\varepsilon_{\hat y} := \varepsilon e_{\hat y}$.
	For any (small enough) $\varepsilon_{\hat y}$, the loss $\inprod{L(\hat \Gamma(p))}{p} \leq \inprod{L(\hat \Gamma(p) + \varepsilon_{\hat y})}{p}$.
\end{lemma}
\begin{proof}
	We denote $L^* := \inprod{L(\hat \Gamma(p))}{p}$.
	We proceed in cases:

	Case 1: $\hat y \in H_p, \varepsilon_{\hat y} > 0$.
	\begin{align*}
	\E_p L(\hat \Gamma(p) + \varepsilon_{\hat y}, Y) &= {\color{blue}\sum_{y \in H_p} p_y (\frac {\varepsilon_{\hat y}}{k})} {\color{teal} - p_{\hat y} \frac{\varepsilon_{\hat y}}{k}} + {\color{blue} \sum_{y \in M_p, B_p} p_y(\frac{\varepsilon_{\hat y}}{k})} + {\color{olive}\sum_{y \in M_p, B_p} p_y(\bar u - \frac 1 k)} + {\color{blue}\sum_{y \in L_p} p_y (\frac{\varepsilon_{\hat y}}{k})}+ {\color{olive}\sum_{y \in L_p}p_y(1+ \bar u)}\\
	&= {\color{blue}\frac{\varepsilon_{\hat y}}{k}} - {\color{teal}p_{\hat y}\frac{\varepsilon_{\hat y}}{k}} + {\color{olive}L^*}
	\end{align*}
	This is greater than or equal to $L^*$ if and only if $\frac{\varepsilon_{\hat y}}{k} \geq p_{\hat y}\frac{\varepsilon_{\hat y}}{k}$.
	This is always true as $p \in \simplex$, so $p_{\hat y} \leq 1$.

@@

	Now, if $H_p = \emptyset$ then we can write the expected loss
	\begin{align*}
	\E_p L(\hat \Gamma(p) - \varepsilon_{\hat y}, Y) &= \sum_{y \in \Y}({\color{olive}\bar u - \frac 1 k} -{\color{blue} \frac{\varepsilon_{\hat y}}{k}}) + {\color{teal} p_{\hat y}\varepsilon_{\hat y}} + \sum_{y \in L_p}({\color{olive}p_y + \bar u - \frac 1 k} -{\color{blue} \frac{\varepsilon_{\hat y}}{k}})\\
	\end{align*}

@@

	First, observe that $H_p = \emptyset \implies B_p = \emptyset$, since we have $i \in B_p \iff p_i \leq \frac{1}{k+1}$ and $p_i > \frac{k(1-p_H)}{(k+1)(k-H)} = \frac 1 {k+1}$, yielding a contradiction.
	\begin{align*}
	p_{\hat y} \varepsilon_{\hat y} \geq \frac{k(1-p_H)}{(k+1)(k-H)} &\geq \frac{1-p_H}{k} &\text{By assumption}\\
	\iff \frac{k}{(k+1)(k-H)} &\geq \frac{1}{k} & \text{Dividing both sides by $(1-p_H)$} \\
	\iff k^2 &\geq (k+1)(k-H) = k^2 - (H-1) k -1 & \text{Manipulating fractions}
	\end{align*}
	The last statement is true since $H \geq 1$, so the result holds.

@@

	\begin{align*}
	\E_p L(\hat \Gamma(p) - \varepsilon_{\hat y}, Y) &= \sum_{y \in M_p, B_p}p_y({\color{olive}\bar u - \frac 1 k}) + \sum_{y \in L_p}p_y({\color{olive}1 + \bar u - \frac 1 k}) + {\color{teal} p_{\hat y}\varepsilon_{\hat y}} \\
	&= {\color{olive}L^*} + {\color{teal} p_{\hat y}\varepsilon_{\hat y}} 
	\end{align*}
	This is greater than or equal to $L^*$ since $\varepsilon_{\hat y} > 0$ (recall instead of adding a negative number, we subtracted a positive one).
\end{proof}

\jessie{I think this is clear, but I'm biased as the writer}
Since $L^{top-k}$ is polyhedral, $\inprod{L^{top-k}(\hat\Gamma(p))}{p}$ is linear in any direction $\varepsilon v$ for small enough $\varepsilon$.
Moreover, $\{\varepsilon_{\hat y}\}_{\hat y \in \Y}$ forms a basis for $\reals^n$, so for all $v \in \reals^n$, there is a unique $\lambda = \{\lambda_i\} \in \reals^n$ such that $v = \lambda_i e_i$.
Therefore, we have 
\begin{align*}
\inprod{L(\hat \Gamma(p) + \sum_i \lambda_{i} \varepsilon_{i})}{p} &= 
\sum_i \lambda_i \inprod{L(\hat \Gamma(p) + \varepsilon_{i})}{p}\\
&\geq \sum_i \lambda_i L^* \\
&= L^*
\end{align*}

@@

However, since $L^{top-k}$ is convex, we actually only need to check vertices of the affine regions on which $u := \hat \Gamma(p)$ sits.  
Moreover, since $L$ is affine on each of those (convex) regions defined by some $T,S$ pair, we actually only need to check the vertices adjacent to $u$.
To justify this, consider that, for a convex region $A^{T,S}$ on which $L^{top-k}$ is affine, let $\hat \Gamma(p)$ have adjacent vertices $a_i$.
For any other vertex $v$ of $A^{T,S}$, there is a point $v_\lambda = \lambda v + (1-\lambda) \hat \Gamma(p)$ for $\lambda \in [0,1]$ such that $v_\lambda = \sum_i \lambda'_i a_i$.\jessie{Right?}

Since $L^{top-k}$ is affine on $A^{T,S}$ and $|haat \Gamma(p)$ attains better loss than its adjacent vertices, we then have 
\begin{align*}
\inprod{p}{L(\hat \Gamma(p))} &\leq \inprod{p}{L(a_i)} \,\, \forall i & \text{by Claim} \\
\implies \inprod{p}{L(\hat \Gamma(p))} &\leq \sum_i \lambda'_i \inprod{p}{L(a_i)} & \text{since $L$ affine} \\
&= \inprod{p}{\sum_i L(\lambda'_i a_i)} & \text{linearity of inner product}\\
&= \inprod{p}{L(\sum_i \lambda'_i a_i)} & \text{affineness}\\
&= \inprod{p}{L(v_\lambda)} & \text{construction of $v_\lambda$}\\
&\leq \inprod{p}{L(v)} & \text{$L$ affine and $v_\lambda$ a convex combo of $v$ and $\hat \Gamma(p)$}\\
\end{align*}

@@

\begin{corollary}
	$\R := \hat \Gamma(\simplex)$ is a (minimum?) finite representative set.
\end{corollary}

@@

We enumerate the vertices of $L^{top-k}$ by first reducing to an infinite representative set $\U$ which allows us to simplify the form of $L^{top-k}$, then define the regions for which $\inprod{p}{L(\cdot)}$ is affine (for any $p$) so that we may enumerate the vertices of $L^{top-k}$.
These vertices have each coordinate ``low'' (e.g., $u_i = 0$), ``medium'' ($u_i = 1$), or ``high'' ($u_i$ is as high as possible while still being in $\U$).
Details are deferred to Appendix~\ref{app:top-k}.

>>>>

	(c) Cannot happen, since $H_p = \emptyset \implies B_p = \left\{i \mid p_i \in \left(\frac{1}{k+1}, \frac{1}{k+1}\right]\right\} = \emptyset$.

	(d) Suppose for contradiction that $H+M > k$, then $M > k-H$.
	Moreover, by simplex constraints, we have \begin{align*}
	\frac{H}{k} + \frac{M}{k+1} < p_H + p_M &\leq 1 & \text{simplex constraints}\\
	\iff (k+1) H + kM &< k^2 + k &\text{distribution}\\
	\iff k(H+M) &< k^2 + k- H &\text{re-arranging}\\
	\implies k &< H & \text{since we assume $k < H+M$}
	\end{align*}
	By construction of $H_p$, this yields a contradiction.

	(e) Suppose $p_H = 1 - \epsilon$.
	For all $b \in B_p$, we then have $p_b > \frac{k \epsilon}{(k+1)(k-H)}$.
	By simplex constraints, we then have 
	\begin{align*}
	1 - \epsilon + \frac{Bk\epsilon}{(k+1)(k-H)}&\leq 1 \\
	\iff \frac{Bk}{(k+1)(k-H)}&\leq 1 \\
	Bk &\leq (k+1)(k-H)\\
	B &\leq \frac{k^2 + k - H - Hk}{k}\\
	B &\leq k+1-H - \frac H k \\
	&< k+1 - H
	\end{align*}
	Thus, $B+H < k+1$.

@@

	(g) It suffices to show $B \leq k - H - M$.
	Suppose $p_H = 1 - \epsilon - \delta$ (with $\epsilon \in [0,1 - \frac H k - \frac M {k+1}]$) and $p_M = \delta$ with $\delta > \frac M {k+1}$.
	We have $p_B \in (\frac{Bk(1-p_H)}{(k+1)(k-H)},\frac{Bk(1-p_H)}{k+1}]$.
	With simplex constraints, we have 
	\begin{align*}
	1 \geq p_H + p_M + p_B > p_H + p_M + \frac{Bk(\epsilon + \delta)}{(k+1)(k-H)}\\
	\iff (1 - \epsilon - \delta) + \delta + \frac{Bk(\epsilon + \delta)}{(k+1)(k-H)} &< 1\\
	\iff \frac{Bk(\epsilon + \delta)}{(k+1)(k-H)} &< \epsilon\\
	\iff B &< \frac{\epsilon (k+1)(k-H)}{k(\epsilon + \delta)}\\
	\implies B &< \frac{\epsilon (k-H)}{\epsilon + \delta}\\
	\implies B &< \frac{\epsilon (k-H)}{\epsilon + \frac M {k+1}}\\
	\implies B &< \frac{1- \frac H k - \frac M {k+1} (k-H)}{1- \frac H k - \frac M {k+1}+ \frac M {k+1}}\\
	\iff B &< \frac{(k- H - \frac {kM}{k+1} (k-H))}{k-H}\\
	\iff B &< k- H - \frac {kM}{k+1}\\
	\implies B &\leq k- H - M\\
	%\implies B&< \frac{\epsilon (k-H)}{\epsilon + \frac M {k+1}}\\
	%\implies B&< \frac{(1 - \frac {H}{k} - \frac{M}{k+1}) (k-H)}{1 - \frac H k}\\
	%\iff B&< \frac{ (k - H - (k+1)M)(k-H)}{k-H}\\
	%\iff B&< k - H - M - kM \\
	%% &< \frac{(k+1)(k-H)}{k}
	%% &< \frac{k^2 + k - Hk - H}{k}\\
	%% &= k + 1 - H - \frac H k
	\end{align*}
	\jessie{TODO}

@@

@@

\begin{lemma}\label{lem:hat-gamma-opt-one-dir}
	Fix $p \in \simplex$.
	For small enough $\varepsilon$ and any $\hat y \in \Y$, consider $\varepsilon_{\hat y} := \varepsilon e_{\hat y}$.
	For any (small enough) $\varepsilon_{\hat y}$, the loss $\inprod{L(\hat \Gamma(p))}{p} \leq \inprod{L(\hat \Gamma(p) + \varepsilon_{\hat y})}{p}$.
\end{lemma}
\begin{proof}
	We denote $L^* := \inprod{L(\hat \Gamma(p))}{p}$ and $\bar u := \bar \sum_{i \in T_k(\hat\Gamma(p))} u_i$.
	We proceed in cases:

	Case 1: $\hat y \in H_p, \varepsilon_{\hat y} > 0$.
	\begin{align*}
	\E_p L(\hat \Gamma(p) + \varepsilon_{\hat y}, Y) &= {\color{blue}\sum_{y \in H_p} p_y (\frac {\varepsilon_{\hat y}}{k})} {\color{teal} - p_{\hat y} \frac{\varepsilon_{\hat y}}{k}} + {\color{blue} \sum_{y \in M_p, B_p} p_y(\frac{\varepsilon_{\hat y}}{k})} + {\color{olive}\sum_{y \in M_p, B_p} p_y(\bar u - \frac 1 k)} + {\color{blue}\sum_{y \in L_p} p_y (\frac{\varepsilon_{\hat y}}{k})}+ {\color{olive}\sum_{y \in L_p}p_y(1+ \bar u)}\\
	&= {\color{blue}\frac{\varepsilon_{\hat y}}{k}} - {\color{teal}p_{\hat y}\frac{\varepsilon_{\hat y}}{k}} + {\color{olive}L^*}
	\end{align*}
	This is greater than or equal to $L^*$ if and only if $\frac{\varepsilon_{\hat y}}{k} \geq p_{\hat y}\frac{\varepsilon_{\hat y}}{k}$.
	This is always true as $p \in \simplex$, so $p_{\hat y} \leq 1$.
	\jessie{Note this case (1) is actually covered by representativeness of $\U$, since we are now leaving $\U$.}

@@

	Now, if $H_p = \emptyset$, then we necessarily have $B_p = \emptyset$ and we can write the expected loss
	\begin{align*}
	\E_p L(\hat \Gamma(p) - \varepsilon_{\hat y}, Y) &=1 + \sum_{y \in \Y}p_y({\color{olive}\bar u - \frac 1 k} -{\color{blue} \frac{\varepsilon_{\hat y}}{k}}) + {\color{teal} p_{\hat y}\varepsilon_{\hat y}} \\%+ \sum_{y \in L_p}({\color{olive}p_y (1 + \bar u - \frac 1 k} -{\color{blue} \frac{\varepsilon_{\hat y}}{k}})\\
	\end{align*}
	Which is greater than or equal to {\color{olive}$L^*$} as $p_{\hat y} \leq \frac 1 k$.

@@

	First, observe that $H_p = \emptyset \implies B_p = \emptyset$, since we have $i \in B_p \iff p_i \leq \frac{1}{k+1}$ and $p_i > \frac{k(1-p_H)}{(k+1)(k-H)} = \frac 1 {k+1}$, yielding a contradiction.
	Now, considering $H_p \neq \emptyset$, we have 
	\begin{align*}
	p_{\hat y} \varepsilon_{\hat y} \geq \frac{k(1-p_H)}{(k+1)(k-H)} &\geq \frac{1-p_H}{k} &\text{By assumption}\\
	\iff \frac{k}{(k+1)(k-H)} &\geq \frac{1}{k} & \text{Dividing both sides by $(1-p_H)$} \\
	\iff k^2 &\geq (k+1)(k-H) = k^2 - (H-1) k -1 & \text{Manipulating fractions}
	\end{align*}
	The last statement is true since $H \geq 1$, so the result holds.

@@

	\begin{align*}
	\E_p L(\hat \Gamma(p) - \varepsilon_{\hat y}, Y) &= \sum_{y \in M_p, B_p}p_y({\color{olive}\bar u - \frac 1 k}) + \sum_{y \in L_p}p_y({\color{olive}1 + \bar u - \frac 1 k}) + {\color{teal} p_{\hat y}\varepsilon_{\hat y}} \\
	&= {\color{olive}L^*} + {\color{teal} p_{\hat y}\varepsilon_{\hat y}} 
	\end{align*}
	This is greater than or equal to $L^*$ since $\varepsilon_{\hat y} > 0$ (recall instead of adding a negative number, we subtracted a positive one).
	\jessie{Again, this would leave $\U$, and is covered by representativeness.}
\end{proof}

@@

However, since $L^{top-k}$ is convex, we actually only need to check vertices of the affine regions on which $u := \hat \Gamma(p)$ sits.  
Moreover, since $L$ is affine on each of those (convex) regions defined by some $T,S$ pair, we actually only need to check the vertices adjacent to $u$.
To justify this, consider that, for a convex region $A^{T,S}$ on which $L^{top-k}$ is affine, let $\hat \Gamma(p)$ have adjacent vertices $a_i$.
For any other vertex $v$ of $A^{T,S}$, there is a point $v_\lambda = \lambda v + (1-\lambda) \hat \Gamma(p)$ for $\lambda \in [0,1]$ such that $v_\lambda = \sum_i \lambda'_i a_i$ (convex combination).\jessie{Right?}

\jessie{Another way to phrase this: for all vertices $v_i$, there is a point $x_i$ in the convex hull of the adjacent vertices such that $v_i$ is in the conic sum of $\hat \Gamma(p)$ and $x_i$}

Since $L^{top-k}$ is affine on $A^{T,S}$ and $\hat \Gamma(p)$ attains better loss than its adjacent vertices, we then have 
\begin{align*}
\inprod{p}{L(\hat \Gamma(p))} &\leq \inprod{p}{L(a_i)} \,\, \forall i & \text{by Lemma~\ref{lem:hat-gamma-opt-one-dir}} \\
\implies \inprod{p}{L(\hat \Gamma(p))} &\leq \sum_i \lambda'_i \inprod{p}{L(a_i)} & \text{since $L$ convex} \\
&= \inprod{p}{\sum_i \lambda'_i L(a_i)} & \text{linearity of inner product}\\
&= \inprod{p}{L(\sum_i \lambda'_i a_i)} & \text{affineness}\\
&= \inprod{p}{L(v_\lambda)} & \text{construction of $v_\lambda$}\\
&\leq \inprod{p}{L(v)} & \text{$L$ affine and $v_\lambda$ a convex combo of $v$ and $\hat \Gamma(p)$}\\
\end{align*}

@@

	Now for $r \in \R$ generated by the sets $H,M$ such that $H, M \subseteq [n]$, $H \cap M = \emptyset$, $|H| + |M| \leq k$, and $|H| < k$ with $k <n$, consider the distribution 
	Fix a small $\varepsilon_M > 0$ and consider $\varepsilon_H = 1 - \frac H k - \frac M {k+1} - M \varepsilon_M$.
	Now consider the distribution p such that, for all $y \in H$, we have $p_y = \frac 1 k + \varepsilon_H$, and similarly, for all $y \in M$, we have $p_y = \frac 1 {k+1} + \varepsilon_M$, and $p_y  = 0 \forall y \in [n] \setminus (H \cup M)$.
	For any $H,M$ combination, we have the generated $p \in \simplex$ \jessie{flesh out that claim}, and therefore $r \in \hat \Gamma(p)$. 
	Thus, $\R \subseteq \hat \Gamma(\simplex)$.

	%$p_{HM} = (\frac{1}{k} + \varepsilon)\ones_{H} + (\frac{1}{k+1} + \varepsilon)\ones_M + (1 - ((\frac{|H|}{k} + |H| \varepsilon) + (\frac{|M|}{k+1} + |M| \varepsilon))) \ones_{[n] \setminus H\cup M} $.
\end{proof}

\begin{corollary}
	$\R$ is a finite representative set.
\end{corollary}

\section{Minimum-ness}
\begin{claim}
	For all $r \in \R$, there is a $p \in \simplex$ such that $\hat \Gamma(p)$ uniquely minimizes $L^{top-k}$.
\end{claim}
\begin{proof}[Sketch]
	Fix $r \in \R$, and consider $H,M$ so that $r$ is generated by $H$ and $M$.

	Since $L^{top-k}$ is affine on the regions $A^{T,S}$ for each $T,S$ affiliated with $H$ and $M$, we just need to show strict inequality in each of the cases of Lemma~\ref{lem:hat-gamma-opt-one-dir}.

	\jessie{Want to formalize, but here's the idea.}
	In Case 1, we would leave $\U$, so no ned to worry.
	In Case 2, we construct $p$ so that $p_{y} < \frac 1 k$, and thus the inequality is strict.
	In Case 3, since $y \in H_p$, we additionally have $p_y > \frac 1 k$.
	In Case 4a, we have $p_y < \frac 1 k$, and case 4b does not happen by construction.
	Case 5 devolves to leaving $\U$, and again the inequality is strict. 
\end{proof}

@@

We enumerate the vertices of $L^{top-k}$ by first reducing to an infinite representative set $\U$ which allows us to simplify the form of $L^{top-k}$, then define the regions for which $\inprod{p}{L(\cdot)}$ is affine (for any $p$) so that we may enumerate the vertices of $L^{top-k}$.
These vertices have each coordinate ``low'' (e.g., $u_i = 0$), ``medium'' ($u_i = 1$), or ``high'' ($u_i$ is as high as possible while still being in $\U$).

Moreover, we characterize a minimizing report of $L^{top-k}$ for each $p \in \simplex$, and show equivalent set of the range of this constructed (single-valued) property is exactly the set $\R$, defined below:

Details are deferred to Appendix~\ref{app:top-k}.
